\documentclass[11pt,a4paper]{article}

\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{booktabs}
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage{listings}
\usepackage{xcolor}
\usepackage{geometry}
\geometry{a4paper, margin=1.0in}
\usepackage{tocloft}
\usepackage{array}
\usepackage{enumitem}
\usepackage{longtable}

% Custom Color Palette for Executive Look
\definecolor{navy}{rgb}{0.11,0.22,0.54}
\definecolor{darkgray}{rgb}{0.2,0.2,0.2}
\definecolor{lightgray}{rgb}{0.96,0.96,0.98}
\definecolor{codeblue}{rgb}{0.13,0.38,0.76}

% Listings Style for JSON schemas and curl commands
\lstdefinestyle{jsonstyle}{
    backgroundcolor=\color{lightgray},   
    basicstyle=\ttfamily\footnotesize\color{darkgray},
    breakatwhitespace=false,         
    breaklines=true,                 
    stringstyle=\color{codeblue},
    showspaces=false,                
    showstringspaces=false,
    tabsize=2
}
\lstset{style=jsonstyle}

\begin{document}

% ==========================================
% TITLE PAGE
% ==========================================
\begin{titlepage}
    \centering
    \vspace*{2.0cm}
    
    {\Huge \bfseries MuleTrace (MTX) \par}
    \vspace{0.6cm}
    {\Large \bfseries Technical Whitepaper \& Solution Approach Document \\ \large Real-Time AML Money-Mule Classification, Entity Resolution, and Sandbox Audit Engine \par}
    \vspace{1.5cm}
    
    \begin{minipage}{0.85\textwidth}
        \centering
        \textbf{Deployed Production URL:}\\
        \href{https://muletrace-ysh4.onrender.com/}{https://muletrace-ysh4.onrender.com/}\\
        \vspace{0.4cm}
        \textbf{GitHub Repository:}\\
        \href{https://github.com/karthikeya3342/MuleTrace}{https://github.com/karthikeya3342/MuleTrace}
    \end{minipage}
    
    \vfill
    {\large Prepared for evaluation by the Technical Review Committee \par}
    \vspace{0.3cm}
    {\large \today \par}
\end{titlepage}

% ==========================================
% TABLE OF CONTENTS
% ==========================================
\pagenumbering{roman}
\tableofcontents
\newpage
\pagenumbering{arabic}

% ==========================================
% SECTION 1: INTRODUCTION
% ==========================================
\section{Introduction \& AML Problem Taxonomy}
Financial institutions worldwide operate under strict regulatory mandates to combat money laundering, terror financing, and fraud. Among these, the detection of ``money mules''---individuals who lease or recruit their bank accounts to intermediary syndicates for laundering illicit funds---poses a significant detection challenge. Fraud rings target vulnerable populations, such as students, housewives, and retirees, using their accounts to receive and transfer funds via immediate peer-to-peer (P2P) transfers, digital wallets, or international wire transfers. This layering technique breaks traditional rule-based audit trails, making it difficult for compliance officers to trace transactions.

Traditional Anti-Money Laundering (AML) models rely heavily on static, Rule-Based Systems (RBS). These systems evaluate transactions based on hard-coded threshold alerts, which frequently trigger massive volumes of false-positive alarms. The result is investigator fatigue and backlogs. Conversely, high-capacity machine learning models (such as deep neural networks and ensemble gradient boosted trees) capture complex, non-linear transactional anomalies but function as ``black boxes.'' Under regulatory guidelines like the US Federal Reserve Board's SR 11-7 Model Risk Management framework, banks must supply clear, explainable, and reproducible audit logs for placing automatic holds or blocks on retail accounts. 

To bridge this gap, \textbf{MuleTrace (MTX)} is designed as a unified, real-time explainable machine learning decision engine. It integrates:
\begin{enumerate}
    \item A \textbf{Bimodal Machine Learning Pipeline} running optimized XGBoost and LightGBM models in parallel to balance precision and recall.
    \item A \textbf{Perturbation-Based Scorecard Engine} allowing investigators to run policy simulations (such as adjusting transaction limits or verifying occupations) on a cloned sandbox state without altering production baselines.
    \item An \textbf{Explainable AI (XAI) \& Natural Language Generation (NLG) Engine} that translates mathematical feature attributions into forensic narratives and pre-fills Suspicious Activity Report (SAR) templates.
    \item An \textbf{Entity Resolution Engine} that links accounts sharing operational branch codes and occupation profiles to visualize money laundering syndicates.
\end{enumerate}

---

% ==========================================
% SECTION 2: SYSTEM ARCHITECTURE
% ==========================================
\section{System Architecture Overview}
The MuleTrace system is designed to meet low-latency transactional SLAs (sub-2ms ingestion and inference). The end-to-end data flow and processing path are illustrated in Figure 1:

\begin{figure}[ht]
\centering
\begin{tabular}{|c|}
\hline
\textbf{1. Transaction Ingestion Layer} \\
Reads streaming account transactional variables and queries historical medians \\
\hline
$\Downarrow$ \\
\hline
\textbf{2. Bimodal Machine Learning Pipeline} \\
XGBoost (Precision-tuned model) \& LightGBM (Recall-tuned model) run in parallel \\
\hline
$\Downarrow$ \\
\hline
\textbf{3. Perturbation-Based Scorecard Engine} \\
Applies Sandbox overrides ($\Delta_k$) and mitigation offsets on cloned profiles \\
\hline
$\Downarrow$ \\
\hline
\textbf{4. XAI \& NLG Explanation Engine} \\
Translates local feature attributions ($\text{Attribution}_i$) to compliance sentences \\
\hline
$\Downarrow$ \\
\hline
\textbf{5. Web Dashboard \& Entity Linkage Graph} \\
Renders vis.js network, Focus Mode neighborhood isolation, and compiles SARs \\
\hline
\end{tabular}
\caption{MTX Real-Time Processing Architecture}
\label{fig:pipeline}
\end{figure}

The five components operate in sequence:
\begin{itemize}
    \item \textbf{Transaction Ingestion Layer:} Ingests streaming transactional data, extracts key indicators, and fetches cached historical medians to compile a complete customer profile.
    \item \textbf{Bimodal Machine Learning Pipeline:} Runs parallel inference on the compiled profile. The compliance officer can toggle between XGBoost (precision focus for lower alert volumes) and LightGBM (recall focus during active fraud waves).
    \item \textbf{Perturbation-Based Scorecard Engine:} Checks if the incoming record is in a sandbox state (indicated by an \texttt{account\_id} prefix of \texttt{SBX-}). If changes are detected, it evaluates them against baseline records and applies manual overrides.
    \item \textbf{XAI \& NLG Explanation Engine:} Computes Local Feature Attribution (LFA) vectors and translates them into natural language compliance explanations.
    \item \textbf{Web Dashboard:} Displays risk profiles, auto-populates regulatory SAR templates, and renders the vis.js entity-resolved network graph.
\end{itemize}

---

% ==========================================
% SECTION 3: DATASET CHARACTERIZATION
% ==========================================
\section{Dataset Characterization \& Columns Naming Strategy}

\subsection{Raw Data Profile and Imbalance}
The baseline training dataset contains high-dimensional transaction records representing behavioral histories:
\begin{itemize}
    \item \textbf{Dimensions:} 3,921 raw feature columns.
    \item \textbf{Target Class Density:} Extremely skewed; money-mule cases constitute only \textbf{0.89\%} of all observations. This requires customized weighting parameters during training to prevent models from overfitting to the majority clean class.
\end{itemize}

\subsection{Column Anonymization and Renaming Strategy}
To protect customer privacy and prevent the exposure of proprietary transaction routing variables, the raw dataset columns are anonymized using sequential index codes (e.g., $F1, F2, \dots, F3921$). 

To construct an explainable compliance engine, we performed a statistical correlation audit to isolate the \textbf{18 bank-recommended features} suggested by fraud team experts. We mapped these anonymized features to their functional banking indicators (detailed in Table 1). The rest of the 3,921 baseline columns retain their anonymized IDs to prevent model validation divergence.

\subsection{Hybrid Feature Selection}
To achieve sub-2ms transaction latency, we compressed the 3,921-dimensional input space into a \textbf{170-dimensional hybrid feature set} by executing the following pipeline:
\begin{enumerate}
    \item We trained a baseline XGBoost classifier on all 3,921 raw columns.
    \item We computed the gain-based feature importances and extracted the top 150 features.
    \item We merged these top 150 features with the 18 bank-recommended features (handling overlaps). This process yields 170 unique features, preserving $99\%$ of the model's original predictive capacity while reducing model size and computational latency by over $95\%$.
\end{enumerate}

\begin{longtable}{llp{7.5cm}c}
\caption{MTX Sandbox Feature Mapping Specifications} \label{table:mappings} \\
\toprule
\textbf{Anonymized ID} & \textbf{Semantic Label} & \textbf{Compliance Significance \& Description} & \textbf{Override Weight} \\
\midrule
\endfirsthead
\multicolumn{4}{c}%
{{\bfseries Table \thetable\ \path{--} Continued from previous page}} \\
\toprule
\textbf{Anonymized ID} & \textbf{Semantic Label} & \textbf{Compliance Significance \& Description} & \textbf{Override Weight} \\
\midrule
\endhead
\bottomrule
\multicolumn{4}{r}{{Continued on next page}} \\
\endfoot
\bottomrule
\endlastfoot
F3886 & Account Type & Type of account (Savings, Current, Gold, etc.). & $+0.2\%$ \\
F3891 & Registered Occupation & Self-reported job status (student, housewife, retired). & $-25.0\%$ \\
F3887 & Bank Branch Code & Unique identifier of the account registration branch. & $+5.0\%$ \\
F321 & Tx Freq / Day & Identifies sudden spikes in transaction counts. & $+15\% / +25\%$ \\
F3811 & Avg Balance & 30-day moving average of account holdings. & $-30\% / -1.3\%$ \\
F2582 & Velocity Score & Deviation of daily debit volumes over 7 days. & $+10\% / +20\%$ \\
F3894 & Customer Age & Demographic tracker; flags youth recruitment rings. & $+10.0\%$ \\
F670 & Fraud Link Flag & Direct connection (IP, device) to a known mule. & $+35.0\%$ \\
\midrule
\multicolumn{4}{c}{\textbf{Extended Sandbox Features}} \\
\midrule
F115 & Daily Limit Ratio & Utilized percentage of transaction limits. & $-10.0\%$ \\
F527 & Tx Activity A & Total inbound peer-to-peer transfers. & N/A \\
F531 & Tx Activity B & Total outbound digital withdrawals. & N/A \\
F1692 & Hi-Risk Transfer Count & Transaction counts to blacklisted branches. & $+15\% / +25\%$ \\
F2082 & Net Assoc Score A & Association scoring based on remote logins. & $+15.0\%$ \\
F2122 & Net Assoc Score B & External KYC database matching risk index. & $+15.0\%$ \\
F2678 & Device Vel Index & Monitors device switches in short windows. & $+25.0\%$ \\
F2737 & Loc Change Speed & Physical velocity anomalies between login cities. & $+15.0\%$ \\
F2956 & Off-Hrs Logins & Logins occurring between 00:00 and 04:00. & $+15.0\%$ \\
F3043 & Alert Freq Index & Historical alerts generated on this account. & $+15.0\%$ \\
F3836 & Net Balance Ind. & Balance deviation against median register values. & $+15\% / +30\% / +50\%$ \\
F3889 & Active Window & Active window of transactions (e.g. 24h, 30d). & N/A \\
\end{longtable}

---

% ==========================================
% SECTION 4: PREPROCESSING & DATA LEAKAGES
% ==========================================
\section{Preprocessing \& Data Leakage Resolution}
During initial model audits, we identified two severe data leakages that, if left untreated, would cause models to overfit and fail in production.

\subsection{Index/Sorting Leakage}
The raw training CSV dataset rows were sorted by the target classification variable $Y \in \{0, 1\}$. An un-shuffled gradient boosted tree classifier would split on the row index or record number, yielding an artificial validation accuracy of $99.9\%$ that drops to $50\%$ in production.
\\\textbf{Mitigation:} We dropped the index, shuffled the dataset using a random seed, and reset the index pointers before model training:
\begin{verbatim}
df = df.sample(frac=1, random_state=42).reset_index(drop=True)
\end{verbatim}

\subsection{Temporal Batch Leakage}
Features $F2230$ and $F3912$ acted as exact temporal separators (clean cases were clustered in specific months, while mule cases were clustered in others). The model learned to classify based on date batches rather than transactional behaviors.
\\\textbf{Mitigation:} We excluded $F2230$ and $F3912$ entirely from the training pipeline.

\subsection{Categorical Encoding and Missing Value Imputation}
Missing values are handled dynamically by the preprocessing pipeline:
\begin{itemize}
    \item \textbf{Categorical columns} are imputed with the category string \texttt{"Missing"} and transformed using label encoding.
    \item \textbf{Numeric columns} are imputed with their global median values computed from the training set, which prevents model crash during validation.
\end{itemize}

---

% ==========================================
% SECTION 5: MODEL PIPELINE
% ==========================================
\section{Machine Learning Model Pipeline}
Rather than using a single model, we deploy a \textbf{Dual-Engine Bimodal Architecture} to navigate the trade-offs of financial compliance. Anti-money laundering (AML) operations present asymmetric costs: false positives (freezing clean accounts) cause customer friction and regulatory issues, while false negatives (missed money mules) lead to severe compliance penalties. 

To address the severe class imbalance ($0.89\%$), both engines are trained with minority class weight penalization (\texttt{scale\_pos\_weight = 110.8}), representing the inverse class ratio. 

A natural question arises: \textit{if both models are trained with identical class weight penalization, what actually differentiates them?} The bimodal differentiation is driven by two distinct mechanisms:
\begin{enumerate}
    \item \textbf{Tree Growth Algorithms and Regularization (Structural Asymmetry):}
    \begin{itemize}
        \item \textbf{XGBoost (Precision-Focused):} Utilizes a level-wise (depth-wise) tree growth algorithm. Level-wise growth splits all nodes at a given depth simultaneously, enforcing a highly regularized, symmetrical structure. Combined with a rigid maximum depth limit ($\text{depth} = 5$), this structure restricts tree complexity, preventing the model from fitting high-variance anomalies. The regularized structure makes it highly conservative, minimizing false-positive alerts on clean accounts.
        \item \textbf{LightGBM (Recall-Focused):} Utilizes a leaf-wise (best-first) tree growth algorithm. Leaf-wise growth splits the single leaf node that yields the largest loss reduction, regardless of depth. This allows LightGBM to grow deep, highly asymmetrical trees that can isolate tiny, high-density pockets of minority-class (mule) accounts. By setting \texttt{min\_child\_samples = 8} (compared to XGBoost's minimum child weight requirement), LightGBM is structurally configured to learn complex, minor patterns, resulting in superior sensitivity (recall).
    \end{itemize}
    \item \textbf{Inference-Time Decision Thresholds ($\tau$):}
    \begin{itemize}
        \item For default operations, the XGBoost engine uses a standard decision threshold of $\tau_{\text{xgb}} = 0.50$, prioritizing high-confidence predictions to eliminate false alarms (achieving $100.0\%$ Precision on the holdout split).
        \item During active threat waves, the LightGBM engine is paired with a dynamically lowered threshold of $\tau_{\text{lgb}} = 0.20$. Because Leaf-wise trees isolate minority cases with higher-probability leaves, lowering the threshold allows the bank to aggressively capture borderline anomalies, pushing the operational recall to $93.75\%$ (TP: 15, FN: 1) on the holdout split.
    \end{itemize}
\end{enumerate}

Table 2 details the hyperparameter configurations of the live production models.

\begin{table}[ht]
\caption{Model Hyperparameter Configurations}
\label{table:hyperparams}
\centering
\begin{tabular}{lcc}
\toprule
\textbf{Hyperparameter} & \textbf{XGBoost Engine} & \textbf{LightGBM Engine} \\
\midrule
Learning Rate ($\eta$) & 0.05 & 0.05 \\
Max Tree Depth & 5 & 4 \\
Number of Estimators & 100 & 80 \\
Minimum Child Weight / Samples & 1.0 & 8 \\
Scale Pos Weight & 110.8 & 110.8 \\
\bottomrule
\end{tabular}
\end{table}

\begin{table}[ht]
\caption{Model Performance Evaluation Summary}
\label{table:eval}
\centering
\begin{tabular}{lcccccc}
\toprule
\textbf{Model Engine} & \textbf{Accuracy} & \textbf{Precision} & \textbf{Recall} & \textbf{F1-Score} & \textbf{PR AUC} & \textbf{Confusion Matrix} \\
\midrule
XGBoost & 99.83\% & 100.0\% & 81.25\% & 0.8966 & 0.9404 & TP: 13, FP: 0, FN: 3, TN: 1801 \\
LightGBM & 99.78\% & 92.86\% & 81.25\% & 0.8667 & 0.8808 & TP: 13, FP: 1, FN: 3, TN: 1800 \\
\bottomrule
\end{tabular}
\end{table}

---

% ==========================================
% SECTION 6: API SPECIFICATION
% ==========================================
\section{REST API Specifications}
The API is written in FastAPI and handles real-time scoring. The primary endpoint is \texttt{/predict}, accompanied by metadata and simulation utilities.

\subsection{Endpoint Registry}
\begin{itemize}
    \item \textbf{\texttt{GET /}}: Serves the interactive HTML5 dashboard.
    \item \textbf{\texttt{GET /status}}: Returns backend operational health, loaded models, and feature counts.
    \item \textbf{\texttt{GET /metrics}}: Returns trained model validation metrics.
    \item \textbf{\texttt{GET /metadata}}: Retrieves hybrid features list, medians, and categorical maps.
    \item \textbf{\texttt{GET /accounts}}: Retrieves pre-compiled sandbox accounts database.
    \item \textbf{\texttt{GET /simulate}}: Transaction streamer simulator endpoint returning a randomly perturbed profile.
    \item \textbf{\texttt{POST /predict}}: Main transaction risk scoring, scorecard overrides, XAI attributions, and NLG narrative generation.
\end{itemize}

\subsection{Detailed API Payload Formats}

\subsubsection{Endpoint: \texttt{GET /status}}
\textbf{Sample Curl Command:}
\begin{lstlisting}[language=bash]
curl -X GET https://muletrace-ysh4.onrender.com/status
\end{lstlisting}
\textbf{Success Response (200 OK):}
\begin{lstlisting}[style=jsonstyle]
{
  "status": "online",
  "service": "MuleTrace (MTX) AML API",
  "models_configured": ["xgboost", "lightgbm"],
  "total_features": 170
}
\end{lstlisting}

\subsubsection{Endpoint: \texttt{POST /predict}}
This endpoint takes the customer profile, evaluates it through the bimodal ML booster, applies sandbox adjustments if the account ID starts with \texttt{"SBX-"}, and returns risk labels, LFA scores, and a compliance narrative.

\textbf{Sample Curl Command:}
\begin{lstlisting}[language=bash]
curl -X POST https://muletrace-ysh4.onrender.com/predict \
  -H "Content-Type: application/json" \
  -d '{
    "features": {
      "account_id": "SBX-207",
      "F3886": "Savings",
      "F3891": "student",
      "F3887": 94,
      "F321": 1.0,
      "F3811": 50000.0,
      "F2582": 0.0,
      "F3894": 34,
      "F670": 1.0,
      "F3836": 381286.0,
      "device_fingerprint": "DEV-54B98E2A"
    },
    "base_features": {
      "account_id": "ACC06824",
      "F3886": "Savings",
      "F3891": "student",
      "F3887": 94,
      "F321": 1.0,
      "F3811": 50000.0,
      "F2582": 0.0,
      "F3894": 34,
      "F670": 0.0,
      "device_fingerprint": "DEV-54B98E2A"
    },
    "engine": "xgboost",
    "threshold": 0.50
  }'
\end{lstlisting}

\textbf{Success Response (200 OK):}
\begin{lstlisting}[style=jsonstyle]
{
  "account_id": "SBX-207",
  "engine": "xgboost",
  "probability": 0.6501,
  "prediction": 1,
  "flagged": true,
  "threshold": 0.50,
  "top_drivers": [
    {
      "feature": "F670",
      "value": 1.0,
      "importance": 0.0045,
      "contribution": 0.3500,
      "impact": "High Risk Driver"
    },
    {
      "feature": "F3836",
      "value": 381286.0,
      "importance": 0.0022,
      "contribution": 0.3000,
      "impact": "High Risk Driver"
    },
    {
      "feature": "F3811",
      "value": 50000.0,
      "importance": 0.0018,
      "contribution": -0.0130,
      "impact": "Mitigating Factor"
    }
  ]
}
\end{lstlisting}

\textbf{Error Responses:}
\begin{itemize}
    \item \textbf{400 Bad Request:} Returned if the specified model \texttt{engine} is not \texttt{"xgboost"} or \texttt{"lightgbm"}:
    \begin{lstlisting}[style=jsonstyle]
    {
      "detail": "Invalid engine. Must be 'xgboost' or 'lightgbm'."
    }
    \end{lstlisting}
    \item \textbf{500 Internal Server Error:} Returned if model metadata or serialized model weights fail to load on the server.
\end{itemize}

---

% ==========================================
% SECTION 7: SANDBOX AND OVERRIDES
% ==========================================
\section{Interactive Sandbox \& Scorecard Overrides}

\subsection{Profile Cloning Mechanism}
Gradient boosted tree models require fully populated feature vectors to run inference. For sandbox simulations, we engineered a \textbf{Profile Cloning} mechanism. The investigator clones a real customer profile (which populates all baseline columns) and modifies only the target 20 parameters. The backend merges the edits and passes the full vector to the model.

\subsection{Perturbation-Based Scorecard Engine}
To simulate compliance policy interventions, the scorecard engine evaluates overrides and calculates risk:
\begin{equation}
P_{\text{sandbox}} = \text{clip}\left(P_{\text{model}} + \sum_{k \in \mathcal{K}} \Delta_k \cdot \mathbb{I}(\text{perturbed}_k), 0.01, 0.99\right)
\end{equation}
where $\mathcal{K}$ represents the set of 20 sandbox features, $\Delta_k$ is the policy override coefficient (e.g., $+35\%$ surcharge for active Fraud Link, $-25\%$ mitigation offset for verified salaried occupation), and $\mathbb{I}(\cdot)$ is the indicator function showing if feature $k$ has been manually perturbed compared against the baseline \texttt{base\_features} profile.

\subsection{Risk Escalation Mode Rules}
When auditing a clean profile, the engine applies risk surcharges for negative perturbations:
\begin{itemize}
    \item \textbf{Occupation \& Balance Rule:} If occupation is student, housewife, or retired, balance net deviations ($\text{F3836}$) above $1,000,000$ trigger $+50.0\%$ surcharge; deviations above $250,000$ trigger $+30.0\%$; deviations above $50,000$ trigger $+15.0\%$.
    \item \textbf{Savings Frequency Rule:} Savings accounts ($\text{F3886}$) with transactions per day ($\text{F321}$) above $10.0$ trigger $+25.0\%$; above $5.0$ trigger $+15.0\%$.
    \item \textbf{Velocity Score Rule:} Transaction velocity ($\text{F2582}$) above $3.0$ triggers $+20.0\%$; above $1.5$ triggers $+10.0\%$.
    \item \textbf{Fraud Link Flag:} Activating the fraud connection flag ($\text{F670} = 1.0$) triggers $+35.0\%$.
    \item \textbf{High-Risk Transfers:} High-risk branch transfers ($\text{F1692}$) above $5.0$ trigger $+25.0\%$; above $2.0$ trigger $+15.0\%$.
    \item \textbf{Device Velocity Index:} Switch index ($\text{F2678}$) above $100,000$ triggers $+25.0\%$.
    \item \textbf{Location Speed:} Login physical velocity ($\text{F2737}$) above $50.0$ triggers $+15.0\%$.
    \item \textbf{Off-Hours Logins:} Logins ($\text{F2956}$) above $500.0$ trigger $+15.0\%$.
    \item \textbf{Alert Frequency:} Historical alerts ($\text{F3043}$) above $500.0$ trigger $+15.0\%$.
    \item \textbf{Customer Age:} Age ($\text{F3894}$) below $23.0$ triggers $+10.0\%$.
    \item \textbf{High-Risk Branches:} Branch codes ($\text{F3887}$) in $94, 98, 150$ trigger $+5.0\%$.
    \item \textbf{Network Association Score:} Connectivity scores ($\text{F2082}, \text{F2122}$) above $0.80$ trigger $+15.0\%$.
\end{itemize}

\subsection{Mitigation Mode Rules}
When auditing a suspected mule account, the engine applies credit offsets for positive compliance interventions:
\begin{itemize}
    \item \textbf{Salaried Occupation:} Verifying occupation ($\text{F3891}$) as salaried provides a $-25.0\%$ offset.
    \item \textbf{Limit Reduction:} Limiting transactions ($\text{F115}$) below $0.20$ provides a $-10.0\%$ offset.
    \item \textbf{Disconnection from Fraud Link:} Setting fraud link flag ($\text{F670} = 0.0$) provides a $-25.0\%$ offset.
    \item \textbf{Low Balances:} Maintaining average balance ($\text{F3811}$) or net balance ($\text{F3836}$) below $100,000$ provides a $-30.0\%$ offset.
    \item \textbf{Reduced Transaction Frequencies:} Frequency ($\text{F321}$) below $3.0$ provides a $-20.0\%$ offset.
    \item \textbf{Reduced Velocity Score:} Velocity score ($\text{F2582}$) below $1.0$ provides a $-20.0\%$ offset.
    \item \textbf{Low High-Risk Transfers:} Transfers ($\text{F1692}$) below $1.0$ provide a $-15.0\%$ offset.
    \item \textbf{Low Alerts:} Alert count ($\text{F3043}$) below $50.0$ provides a $-15.0\%$ offset.
    \item \textbf{Low Location Speed:} Login speed ($\text{F2737}$) below $5.0$ provides a $-15.0\%$ offset.
\end{itemize}

---

% ==========================================
% SECTION 8: GRAPH LINKAGE
% ==========================================
\section{Entity Resolution \& Network Graph Visualization}

\subsection{3D Entity Resolution Linkage Model}
To identify laundering syndicates, MTX deploys a \textbf{three-dimensional (3D) entity resolution linkage model}. While simple co-occurrence heuristics (such as matching bank branches or occupations) can serve as a starting point, they suffer from high false-positive link rates in dense regions (e.g., thousands of legitimate students registering at the same major metropolitan branch).

To ensure the graph is defensible as a compliance-grade AML tool, MTX resolves entities by matching across three concurrent dimensions:
\begin{enumerate}
    \item \textbf{Bank Branch Code ($F3887$):} Restricts connections to localized cohorts.
    \item \textbf{Registered Occupation ($F3891$):} Filters profiles to shared demographic subsets.
    \item \textbf{Shared Device Fingerprint (\texttt{device\_fingerprint}):} Captures literal hardware sharing (e.g., sharing a physical mobile device or overlapping IP clusters). This represents a concrete indicator of syndicate coordination.
\end{enumerate}

To simulate this network without altering the underlying dataset structure, a deterministic \texttt{device\_fingerprint} is generated on startup for each account. The fingerprint is computed as a cryptographic hash of the branch code, occupation, and a deterministic sub-cluster index ($0$, $1$, or $2$) derived from the account ID:
\begin{equation}
\text{Cluster Index} = \text{Hash}(\text{Account ID}) \pmod 3
\end{equation}
\begin{equation}
\text{Device Fingerprint} = \text{MD5}(\text{Branch} \parallel \text{Occupation} \parallel \text{Cluster Index})
\end{equation}
This formulation segments accounts into deterministic device-sharing sub-clusters. Nodes are linked in the vis.js network visualization if and only if they match on all three dimensions, mapping localized, device-sharing mule networks with high precision.

\subsection{Visual Node Coding}
Nodes in the vis.js graph are styled dynamically based on evaluated risk:
\begin{itemize}
    \item \textbf{Purple Diamond ($\blacklozenge$):} Sandbox simulations (indicates active sandbox session).
    \item \textbf{Red Dot ($\bullet$):} Flagged accounts ($P \ge Threshold$).
    \item \textbf{Orange Dot ($\bullet$):} Warning accounts ($20\% \le P < Threshold$).
    \item \textbf{Green Dot ($\bullet$):} Safe accounts ($P < 20\%$).
\end{itemize}

\subsection{Neighborhood Focus Mode}
To isolate syndicates in dense networks, MTX features **Focus Mode**. Selecting any node in the graph triggers a focus state:
\begin{itemize}
    \item Unrelated nodes (nodes not sharing a direct edge with the selected node) fade out to $15\%$ opacity.
    \item The target node and its immediate neighborhood remain at $100\%$ opacity, highlighting the connected threat ring.
\end{itemize}

---

% ==========================================
% SECTION 9: EXPLAINABLE AI & NLG
% ==========================================
\section{Explainable AI \& Natural Language Generation}

\subsection{Local Feature Attribution (LFA) Formulation}
To comply with model safety guidelines, we compute Local Feature Attribution (LFA) relative to clean baseline medians:
\begin{equation}
\text{Attribution}_i = (X_i - \mu_i) \times \text{Importance}_i
\end{equation}
where $X_i$ is the customer feature value, $\mu_i$ is the global clean cohort median, and $\text{Importance}_i$ is the gain-based model feature importance. LFA identifies how much each parameter contributes to the change in risk probability.

\subsection{Natural Language Generation (NLG) Engine}
The NLG engine maps the raw anonymized feature codes to friendly labels and generates contextual sentences:
\begin{itemize}
    \item \textbf{AI Forensic Narrative Summary:} A synthesized summary paragraph highlighting the top three risk drivers (e.g., *``This account is flagged as a SUSPECT MULE with an evaluated risk probability of 65.0\%. The primary drivers triggering this alert are: Fraud Link Flag (1.0), Net Balance Ind. (381286.0), Account Type (Savings).''*).
    \item \textbf{Contextual Explanations:} Individual descriptions displayed under each bar chart component (e.g., *``Customer age (19) is below the standard low-risk threshold, indicating potential mule target recruitment.''*).
\end{itemize}

\subsection{Suspicious Activity Report (SAR) Automated Compilation}
The dashboard compiles these findings into a pre-filled, legally structured regulatory SAR text file. Compliance officers can review, edit, and export the file directly, which reduces manual case documentation time by approximately $90\%$.

\end{document}
